Consistent Polyhedral Surrogates via Embedding

### 2019-07-29 notes/top-k.tex
Simplified argument without cases

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

@@

\subsection{Assumptions and notation}
\subsubsection{Assumptions}
Throughout, we make two assumptions for ease of exposition, but both follow without loss of generality for an optimal report.
First, assume that $u$ is ordered, where $u_1 \geq u_2 \geq \ldots \geq u_n$.
This is without loss of generality, as we can simply reorder $p$ if this is not the case.

@@

\subsubsection{Notation}
Let $\bar{u} = \frac 1 k \sum_{i = 1}^k u_i$ and similarly, $\bar{u}_{-y} = \frac{1}{k-1} \sum_{i=1, i \neq y}^k u_i$ be the averages of the first $k$ elements of $u$ and the average of the first $k$ elements of $u$ besides $y$, respectively.

We use $\odot$ to denote the Hadamard (element-wise) vector product.

@@

First, observe that we can re-write
\begin{align*} 
L^{top-k}(u)_y&=\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+\\ 
&=\left(1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+\\
&=\left(1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - \min(u_y, 1)\Ind{y \leq k})_{i} \right)_+\\
\end{align*}

First, we break this into three cases:
\begin{enumerate}
	\item [$y > k$] Here, we observe that $L^{top-k}(u)_y = (1-u_y + \bar{u})_+ > 0$ for all $u$ given our assumptions, so we don't need to worry about the $\max(\cdot,0)$ operation.
	\item [$y \leq k$, $u_y \leq 1$] Here, $\min(u_y,1) = u_y$, so we can rewrite the loss as $L^{top-k}(u)_y = (1 - \frac {k+1}{k} u_y + \bar u)_+$, which always has the inside term nonnegative.
	\item [$y \leq k$, $u_y > 1$] The term inside $\cdot_+$ of $L^{top-k}(u)_y$ is nonnegative and equal to $1 - \frac 1 k - u_y + \ubar$ when $u_y \leq \bar{u}_{-y} + 1$, and $0$ if $u_y \geq \bar{u}_{-y} + 1$.
\end{enumerate}

Therefore, the expected loss can by written as four summations.
We introduce the constants $t < b \leq k$ to mark the boundaries between cases.
We say $t = \argmax_i \{u_i : u_i > \bar u_{-i} +1\}$ and $b = \argmax_i \{u_i : u_i > 1\}$.
We will argue that we can assume $t = 0$ (i.e. there is no $u_i$ such that $u_i > \bar u_{-i} +1$) for a minimizing report, since reassigning such a $u'_i := \bar u_{-i} +1$ results in $L^{top-k}(u')_i = L^{top-k}(u)_i = 0$, but reduces $L^{top-k}(\cdot)_j$ for all $j \neq i$ by decreasing $\ubar$.

We can now write the expected loss as follows:
\begin{align*}
\inprod{p}{L^{top-k}(u)} &= \sum_{y=1}^n p_y L^{top-k}(u)_y \\
&= \sum_{y=1}^t p_y L^{top-k}(u)_y + \sum_{y=t+1}^b p_y L^{top-k}(u)_y + \sum_{y=b+1}^k p_y L^{top-k}(u)_y + \sum_{y=k+1}^n p_y L^{top-k}(u)_y\\
&= 0 + \sum_{y=t+1}^b p_y (1 - \frac 1 k - u_y + \ubar) + \sum_{y=b+1}^k p_y (1 - \frac{k+1} k u_y + \ubar) + \sum_{y=k+1}^n p_y(1-u_y + \ubar)\\
\end{align*}
Moreover, substituting $j=0$, we can simplify further:
\begin{align*}
&= 1 + \ubar - \inprod{p}{u} - \inprod{p \odot \Ind{y \leq i}}{\frac 1 k \ones} - \inprod{p \odot \Ind{i+1 \leq y \leq k}}{ \frac u k}\\
&= 1 + \ubar - \inprod{p \odot \Ind{y \leq i}}{\frac 1 k \ones + u} - \inprod{p \odot \Ind{i+1 \leq y \leq k}}{ \frac {k+1} k u}\\
\end{align*}
When considering the Bayes Risk, we use this final form of the expected loss.

\subsection{Bayes Risk}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

@@

\subsection{Notation and Assumptions}
\subsubsection{Notation}
Let $\bar{u} = \frac 1 k \sum_{i = 1}^k u_i$ and similarly, $\bar{u}_{-y} = \frac{1}{k-1} \sum_{i=1, i \neq y}^k u_i$ be the averages of the first $k$ elements of $u$ and the average of the first $k$ elements of $u$ besides $y$, respectively.
Moreover, we say $b = \argmax_i \{u_i : u_i > 1\}$.

We use $\odot$ to denote the Hadamard (element-wise) vector product.
Also, $\min(\vec u, \vec v)$ is the element-wise minimum of vectors $\vec u, \vec v \in \reals^d$.

\subsubsection{Assumptions}
Throughout, we make two assumptions for ease of exposition, but both follow without loss of generality for an optimal report.
First, assume that $u$ is ordered, where $u_1 \geq u_2 \geq \ldots \geq u_n$.
This is without loss of generality, as we can simply reorder $p$ if this is not the case.

@@

Observe that, for a minimizer of $L^{top-k}$, we can assume there is no $u_i$ such that $u_i > \bar u_{-i} +1$, since reassigning such a $u'_i := \bar u_{-i} +1$ results in $L^{top-k}(u')_i = L^{top-k}(u)_i = 0$, but reduces $L^{top-k}(\cdot)_j$ for all $j \neq i$ by decreasing $\ubar$.

@@

First, observe that we can re-write
\begin{align*} 
L^{top-k}(u)_y&=\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+\\ 
&=\left(1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - e_y)_{[i]} \right)_+\\
&=\left(1 - u_y + \frac{1}{k} \sum_{i=1}^k (u - \min(u_y, 1)\Ind{y \leq k})_{i} \right)_+
\end{align*}

One can verify that, given our assumptions on $u$, the term inside $(\cdot)_+$ of $L^{top-k}(u)_y$ is nonnegative, so we can drop the $(\cdot)_+$ operator on this report space.

We can now write the expected loss as follows:
\begin{align*}
\inprod{p}{L^{top-k}(u)} &= \sum_{y=1}^n p_y L^{top-k}(u)_y \\
&= \sum_{y=1}^b p_y L^{top-k}(u)_y + \sum_{y=b+1}^k p_y L^{top-k}(u)_y + \sum_{y=k+1}^n p_y L^{top-k}(u)_y\\
&= \sum_{y=1}^b p_y (1 - \frac 1 k - u_y + \ubar) + \sum_{y=b+1}^k p_y (1 - \frac{k+1} k u_y + \ubar) + \sum_{y=k+1}^n p_y(1-u_y + \ubar)\\
&= 1 + \ubar - \inprod{p}{u} - \inprod{p \odot \Ind{y \leq i}}{\frac 1 k \ones} - \inprod{p \odot \Ind{i+1 \leq y \leq k}}{ \frac u k}\\
&= 1 + \ubar - \inprod{p}{u + \frac 1 k \min(u, \ones)}
\end{align*}
When considering the Bayes Risk, we use this final form of the expected loss.

\subsection{Bayes Risk}
First, consider the case where $u_y = 0$ and $u'_y = 1$, with $u_j = u'_j \;\forall i \neq j$.
As the expected loss is linear on $u_y$, and these are the boundary cases, we know that $u$ is optimal on one of these two reports in this case.

So we ask when is $\inprod{p}{L^{top-k}(u)} \leq \inprod{p}{L^{top-k}(u')}$?
\begin{align*}
1 + \ubar - \inprod{p}{u + \frac 1 k \min(u, \ones)} &\leq 1 + \ubar' - \inprod{p}{u' + \frac 1 k \min(u', \ones)}\\
- \inprod{p}{u + \frac 1 k \min(u, \ones)} &\leq \frac {1}{k} - \inprod{p}{u' + \frac 1 k \min(u', \ones)}\\
0 &\leq \frac{1}{k} - p_y(\frac{k+1}{k})\\
p_y &\leq \frac{1}{k+1}
\end{align*}

Conversely, if $p_y > \frac{1}{k+1}$, it is to better to report $u_y = 1$ than $u_y = 0$.
However, since we are not restricted to the unit hypercube, we might ask when it is optimal to bring $u_y = 1$ up to $u'_y = 1 + \epsilon$ given $\epsilon > 0$.
Here, we consider when $\inprod{p}{L^{top-k}(u)} \leq \inprod{p}{L^{top-k}(u')}$.
\begin{align*}
1 + \ubar - \inprod{p}{u + \frac 1 k \min(u, \ones)} &\leq 1 + \ubar' - \inprod{p}{u' + \frac 1 k \min(u', \ones)}\\
\frac{1}{k} - p_y (\frac{k+1}{k}) &\leq \frac{1+\epsilon}{k} - p_y (\frac{k+1}{k}(1 + \epsilon))\\
p_y &\leq \frac{1}{k}
\end{align*}
In this case, $u_y=1$ is preferable to $u+y >1$.
If $p_y > \frac 1 k$, then we want to push $u_y$ as high as we can, but recall that we restrict to $u_y \leq \ubar_{-y} + 1$.
Because the upper bound on $u_y$ depends on $\ubar$, we can't determine $u$ iteratively.
However, we can construct $u$ as given below.
\begin{conjecture}
	Fix $p \in \simplex$.
	Let $g_t = |\{p_i : p_i \geq \frac 1 k\}|$ and $g_b = | \{p_i \in (\frac{1}{k+1}, \frac{1}{k})\}|$.
	The property \[\gamma(p) = 
	\left[
	\begin{cases}
	0 & p_i \leq \frac {1}{k+1}\\
	1 & p_i \in (\frac{1}{k+1}, \frac{1}{k})\\
	\frac{g_b + k-1}{k-g_t} & o/w
	\end{cases}
	\right]_{i=1}^n
	\] refines the property $\prop{L^{top-k}}$.
\end{conjecture}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
